---
title: "Analysis of ToothGrowth data"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

* *Author - Sanjeev Kumar Maheve*
* *Dated -  January 27, 2016*

### Overview
The report aims to analyze the ToothGrowth data in the R datasets package.

```{r setoptions,echo=TRUE}
library(knitr)
# Setting the global options.
opts_chunk$set(fig.width=6, fig.height=4, warning=FALSE)
```

### Load the ToothGrowth data and perform some basic exploratory data analyses.

```{r}
library(datasets)
data(ToothGrowth)
# Print the frame details
str(ToothGrowth)

unique(ToothGrowth$len); unique(ToothGrowth$dose)

library(ggplot2)
m <- ggplot(ToothGrowth, aes(x=factor(dose), y=len))
m + geom_boxplot() + facet_grid(.~supp) + ggtitle("ToothGrowth data details")
```

### Provide a basic summary of the data.
Included figures highlight the means we are comparing.
```{r}
summary(ToothGrowth)
```

### Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
```{r}
# TBD
```

### State your conclusions and the assumptions needed for your conclusions
Based on the visual inspection of data from the box plot; there is a direct relationship between the size of the tooth and the dose required for the same.
